Dynamics of isoflurane-induced vasodilation and blood flow of cerebral vasculature revealed by multi-exposure speckle imaging

### 2021-10-28 article.tex
Response to Reviewer #3

The use of general anesthesia during neuroimaging is ubiquitous across many animal models despite systemic effects on neurophysiological state and cardiovascular function \cite{Slupe:2018ea}. Volatile inhalation anesthetics, such as the halogenated ether isoflurane \cite{Eger:1981ia}, are commonly utilized to immobilize animals during imaging while allowing for fine control over the depth of anesthesia and consciousness. However, isoflurane has been shown to reduce neuronal activity \cite{Aksenov:2015ea} and functional connectivity \cite{Xie:2019de}, suppress the magnitude and speed of neurovascular coupling \cite{Masamoto:2012bj,Takuwa:2012ee,Pisauro:2013cx}, and induce significant vasodilation \cite{Koenig:1994rn,Iida:1998th}. Isoflurane also conveys potential neuroprotective effects that reduce and delay the severity of cerebral ischemia \cite{Kitano:2007ia,Kawaguchi:2000id,Sakai:2007wc,Li:2013ip,Lu:2017bo}. These effects can mask the benefits of prospective neuroprotective therapeutics or interventions and confound the outcomes of long-term studies \cite{Kapinya:2002ua,Seto:2014ga}. For these reasons and more \cite{Gao:2017tw}, there is a growing effort within the neuroscience community to transition to un-anesthetized awake animal models in order to more accurately interpret neurophysiological and behavioral experiments and more readily translate findings to awake human physiology.

@@

Mice ($n = 4$, C57BL/6J, male, 4-6 months, Charles River Laboratories, Inc.) were anesthetized with medical \ce{O2} vaporized isoflurane (3\% for induction, 1-2\% for maintenance, 0.5 L/min) via nose-cone inhalation. Body temperature was maintained at 37 $^\circ$C with a feedback heating pad (DC Temperature Controller, Future Health Concepts, Inc.). Arterial oxygen saturation, heart rate, and breathing rate were monitored via pulse oximetry (MouseSTAT, Kent Scientific Corp.). Mice were placed supine in a stereotaxic frame (Kent Scientific Corp.) and subcutaneously administered carprofen (5 mg/kg) and dexamethasone (2 mg/kg) to reduce inflammation during the craniotomy procedure. The scalp was shaved and resected to expose the skull between the bregma and lambda cranial coordinates. A circular region (4 mm diameter) of the skull including the dura mater over the somatosensory cortex was removed with a dental drill (Ideal Microdrill, 0.8 mm burr, Fine Science Tools, Inc.) while under regular artificial cerebrospinal fluid (buffered pH 7.4) perfusion. A 4 mm round \#1.5 cover glass (World Precision Instruments, Inc.) was placed over the exposed brain with artificial cerebrospinal fluid filling the void. Gentle pressure was applied to the cover glass while Kwik-Sil adhesive (World Precision Instruments, Inc.) was deposited to bond the glass to the skull. A layer of Vetbond tissue adhesive (3M) was applied over the Kwik-Sil to create a sterile, air-tight seal around the craniotomy and to allow for the restoration of intracranial pressure. C\&B Metabond (Parkell, Inc.) was then used to cement a custom titanium head-plate centered around the cranial window for head-constrained awake measurements. The subjects were allowed to recover from surgery and monitored for cranial window integrity and normal behavior for at least four weeks prior to imaging. They were then exposed to head fixation during 20-30 minute sessions over 3-5 days until habituated to locomotion on a linear treadmill awake imaging system \cite{He:2020}.

@@

Cranial window implanted mice were head constrained on the awake imaging system and allowed to acclimate for 5-10 minutes. The awake hemodynamic baseline was then defined by acquiring 50 MESI frames, averaging the speckle contrast images across each exposure time, and computing the corresponding MESI $ICT$ image using Eq.~(\ref{eq:mesi}). A custom 3D-printed inhalation nose-cone connected to the anesthesia vaporizer and scavenger was then placed several millimeters away from the subject and used to administer medical air (0.5 L/min) with no isoflurane (0\%). Continuous MESI was initiated ($t = 0$) and used to monitor the awake subject for 10 minutes before increasing isoflurane to 2.0\% to induce anesthesia. After one minute of exposure ($t = 11$ minutes), the nose-cone was fully positioned over the subject to ensure targeted delivery of isoflurane. After two minutes of exposure ($t = 12$ minutes), a feedback heating pad (55-7030, Harvard Apparatus, Inc.) was placed beneath the subject to maintain a 37 $^\circ$C body temperature. The pulse oximeter was also attached and used to monitor vitals and regulate the depth of anesthesia such that the heart rate was matched to within 10\% between imaging sessions for each subject \cite{Kazmi:2013hp}. The now anesthetized subject was continuously monitored with MESI until $t = 30$ minutes, at which point the anesthetized hemodynamic state was defined by acquiring an additional 50 MESI frames. The speckle contrast images were then averaged across each exposure time and used to compute the MESI $ICT$ image using Eq.~(\ref{eq:mesi}). The subject was then removed from the imaging system and allowed to recover from anesthesia on a heating pad before being transferred back into its cage. Each subject underwent three imaging sessions with at least one full day of recovery between each session.

@@

While image registration can help maintain the spatial alignment of data across an experiment, it is unable to decouple the effects of animal motion from the underlying blood flow. Even with the head fully restrained, walking and grooming both caused subtle brain movements that manifested as large changes in speckle contrast, as seen by the abrupt spikes during the awake section of Fig.~\ref{fig:timecourse}c. While these fluctuations may represent real hemodynamic responses, it is difficult to isolate them from broader animal motion. A simple solution might be the addition of an external sensor to the treadmill to exclude timepoints when the animal is actively moving \cite{Dombeck:2007gr}. However, a more refined awake imaging system that further minimizes brain movement would likely be necessary to directly probe these phenomena with LSCI or MESI.

@@

We have used MESI to continuously image CBF during the induction of general anesthesia with isoflurane in head-restrained mice. The vasodilatory effects of isoflurane caused rapid and large anatomical and physiological changes that were drastically different from the awake state. These increases in vessel size and blood flow were repeatable across multiple imaging sessions and subjects. We also documented disparate response times to the induction of isoflurane anesthesia between arteries, veins, and parenchyma regions. This study demonstrated that MESI can be readily used for chronic awake imaging and allows for direct day-to-day comparisons of blood flow. These results provide further evidence that the anesthetized state is unrepresentative of normal physiology and that neurophysiological and behavioral experiments would benefit immensely from transitioning to un-anesthetized awake animal models.

>>>>

The use of general anesthesia during neuroimaging is ubiquitous across many animal models despite systemic effects on neurophysiological state and cardiovascular function \cite{Slupe:2018ea}. Volatile inhalation anesthetics, such as the halogenated ether isoflurane \cite{Eger:1981ia}, are commonly utilized to immobilize animals during imaging while allowing for fine control over the depth of anesthesia and consciousness. However, isoflurane has been shown to reduce neuronal activity \cite{Aksenov:2015ea}, interfere with functional connectivity at higher doses \cite{Xie:2019de}, and suppress the magnitude and speed of neurovascular coupling \cite{Masamoto:2012bj,Takuwa:2012ee,Pisauro:2013cx}. Isoflurane also conveys potential neuroprotective effects that reduce and delay the severity of cerebral ischemia \cite{Kitano:2007ia,Kawaguchi:2000id,Sakai:2007wc,Li:2013ip,Lu:2017bo}. These effects can mask the benefits of prospective neuroprotective therapeutics or interventions and confound the outcomes of long-term studies \cite{Kapinya:2002ua,Seto:2014ga}. For these reasons and more \cite{Gao:2017tw}, there is a growing effort within the neuroscience community to transition to un-anesthetized awake animal models in order to more accurately interpret neurophysiological and behavioral experiments and more readily translate findings to awake human physiology.

@@

Mice ($n = 4$, C57BL/6J, male, 4-6 months, Charles River Laboratories, Inc.) were anesthetized with medical \ce{O2} vaporized isoflurane (3\% for induction, 1-2\% for maintenance, 0.5 L/min) via nose-cone inhalation. Body temperature was maintained at 37 $^\circ$C with a feedback heating pad (DC Temperature Controller, Future Health Concepts, Inc.). Arterial oxygen saturation, heart rate, and breathing rate were monitored via pulse oximetry (MouseSTAT, Kent Scientific Corp.). Mice were placed supine in a stereotaxic frame (Kent Scientific Corp.) and subcutaneously administered carprofen (5 mg/kg) and dexamethasone (2 mg/kg) to reduce inflammation during the craniotomy procedure. The scalp was shaved and resected to expose the skull between the bregma and lambda cranial coordinates. A circular region (4 mm diameter) of the skull including the dura mater over the somatosensory cortex was removed with a dental drill (Ideal Microdrill, 0.8 mm burr, Fine Science Tools, Inc.) while under regular artificial cerebrospinal fluid (buffered pH 7.4) perfusion. A 4 mm round \#1.5 cover glass (World Precision Instruments, Inc.) was placed over the exposed brain with artificial cerebrospinal fluid filling the void. Gentle pressure was applied to the cover glass while Kwik-Sil adhesive (World Precision Instruments, Inc.) was deposited to bond the glass to the skull. A layer of Vetbond tissue adhesive (3M) was applied over the Kwik-Sil to create a sterile, air-tight seal around the craniotomy and to allow for the restoration of intracranial pressure. C\&B Metabond (Parkell, Inc.) was then used to cement a custom titanium head-plate centered around the cranial window for head-constrained awake measurements. The subjects were allowed to recover from surgery and monitored for cranial window integrity and normal behavior for at least four weeks prior to imaging. They were then exposed to head fixation during 20-30 minute sessions over 3-5 days until habituated to locomotion on a custom, low-profile linear treadmill system \cite{He:2020}.

@@

Cranial window implanted mice were head constrained on the awake imaging system and allowed to acclimate for 5-10 minutes. Periods of extended locomotion on the treadmill were manually noted throughout the experiment. The awake hemodynamic baseline was defined by acquiring 50 MESI frames, averaging the speckle contrast images across each exposure time, and computing the corresponding MESI $ICT$ image using Eq.~(\ref{eq:mesi}). A custom 3D-printed inhalation nose-cone connected to the anesthesia vaporizer and scavenger was then placed several millimeters away from the subject and used to administer medical air (0.5 L/min) with no isoflurane (0\%). Continuous MESI was initiated ($t = 0$) and used to monitor the awake subject for 10 minutes before increasing isoflurane to 2.0\% to induce anesthesia. After one minute of exposure ($t = 11$ minutes), the nose-cone was fully positioned over the subject to ensure targeted delivery of isoflurane. After two minutes of exposure ($t = 12$ minutes), a feedback heating pad (55-7030, Harvard Apparatus, Inc.) was placed beneath the subject to maintain a 37 $^\circ$C body temperature. A pulse oximeter (MouseSTAT, Kent Scientific Corp.) was also attached and used to monitor vitals and regulate the depth of anesthesia such that the heart rate was matched to within 10\% between imaging sessions for each subject \cite{Kazmi:2013hp}. The now anesthetized subject was continuously monitored with MESI until $t = 30$ minutes, at which point the anesthetized hemodynamic state was defined by acquiring an additional 50 MESI frames. The speckle contrast images were then averaged across each exposure time and used to compute the MESI $ICT$ image using Eq.~(\ref{eq:mesi}). The subject was then removed from the imaging system and allowed to recover from anesthesia on a heating pad before being transferred back into its cage. Each subject underwent three imaging sessions with at least one full day of recovery between each session.

@@

While image registration can help maintain the spatial alignment of data across an experiment, it is unable to decouple the effects of animal motion from the underlying blood flow. Even with the head fully restrained, walking and grooming both caused subtle brain movements that manifested as large changes in speckle contrast, as seen by the abrupt spikes during the awake section of Fig.~\ref{fig:timecourse}c. While these fluctuations may represent real hemodynamic responses, it is difficult to isolate them from broader animal motion given LSCI's inherent sensitivity to any movement \cite{Richards:2014kx}. A simple solution might be the addition of an external sensor to the treadmill to automatically exclude timepoints when the animal is actively moving \cite{Dombeck:2007gr}. Alternatively, a webcam could be used to monitor all animal activity, which could facilitate comparisons between animals at rest, grooming, or locomoting. However, a more refined awake imaging system that further minimizes brain movement may likely be necessary to directly probe these phenomena with LSCI or MESI.

@@

We have used MESI to continuously image CBF during the induction of general anesthesia with isoflurane in head-restrained mice. The vasodilatory effects of isoflurane caused rapid and large anatomical and physiological changes that were drastically different from the awake state. These increases in vessel size and blood flow were repeatable across multiple imaging sessions and subjects. We also documented disparate response times to the induction of isoflurane anesthesia between arteries, veins, and parenchyma regions. This study demonstrated that MESI can be readily used for chronic awake imaging and allows for direct day-to-day comparisons of blood flow. There are numerous applications for the longitudinal characterization of CBF dynamics in studying the progression of ischemic stroke absent the effects of anesthesia and the efficacy of potential therapeutics targeting flow restoration. These results also provide further evidence that the anesthetized state is unrepresentative of normal physiology and that neurophysiological and behavioral experiments would benefit immensely from transitioning to un-anesthetized awake animal models.
